---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

library(emoji)
```

# Consultation, Co-Optation, and Crackdown: How Participation in Authoritarian-Dominated International Organizations Shapes Civil Society

[Andrew Heiss](https://www.andrewheiss.com) • Georgia State University

<!-- badges: start -->
<!-- badges: end -->

---

## Abstract

A growing body of research currently demonstrates that state membership in authoritarian-dominated regional and international organizations (ROs and IOs) can weaken the prospects for political liberalization and can serve to bolster autocratic regime stability. Concurrently, over the past decade, the regulatory environment for nongovernmental organizations (NGOs) has become increasingly restricted, particularly in authoritarian regimes where autocrats wield anti-civil society regulations in order to maintain and increase regime stability. NGOs play an important role in international policymaking, often holding consultative or voting roles in various types of IOs. Given the increasing restriction of civil society in authoritarian states, what role do NGOs play in authoritarian-oriented IOs? In this paper, I explore two questions related to authoritarian stability, civil society, and international organizations: (1) how does membership in authoritarian IOs shape states’ domestic civil society regulations and NGO behavior, and (2) how does NGO participation in IO deliberations differ from that in democratic-oriented IOs? I explore these questions with cross-national quantitative data from the Varieties of Democracy project and two qualitative case studies of NGOs working with authoritarian IOs.


### `r emoji_glue(":mage::worm:")`: Note on "experienced earthworm" project name

Because project titles change all the time with revisions, rewriting, and peer review, we used [{codename}](http://svmiller.com/codename/) to generate an Ubuntu-style internal-to-me project name that won't change.

```{r generate-code-name, eval=FALSE}
library(codename)
codename_message()
#> code name generated by {codename} v.0.5.0. R version 4.3.0 (2023-04-21).

codename(type = "ubuntu", seed = "authn IOs and NGOs")
#> [1] "experienced earthworm"
```


## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).

## Contributions and Code of Conduct

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
